---
title: "Admissions Data - Cleaning Test Scores/GPA"
author: "Lydia LaSeur"
date: "4/15/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(data.table)
library(tidyverse)
library(kableExtra)
```


Read in raw data for admitted students and select only the columns we need.  

```{r}
# both admitted.csv and denied.csv are the raw data from Google sheets exported as a csv
admit_raw <- fread("./PrivateData/admitted.csv", sep = ",")
str(admit_raw)


admit <- admit_raw %>%
  select(Ref, Round, `Course Selection Rating`, `Final NCF GPA`, `ACT Composite`, `SAT I Superscore (CR + M) - Verified`, `SAT R Superscore`) %>%
  mutate(Accepted = 1)

colnames(admit) <- c("Ref", "Round", "Rigor","GPA","ACT","SAT_I", "SAT_R","Accepted")
```


Do the same for the raw data for the denied students.  

```{r}
denied_raw <- fread("./PrivateData/denied.csv", sep = ",")
str(denied_raw)

# there are multiple cols for GPA, merging them to 1 column
all_gpa_cols <- denied_raw[,c(10,11,12,13)]
all_gpa_cols[is.na(all_gpa_cols)] <- 0
denied_GPA <- rowSums(all_gpa_cols)

denied <- denied_raw
denied <- cbind(denied, GPA = denied_GPA)

denied <- denied %>%
  select(Ref, Round, `Course Selection Rating`, GPA, `ACT Composite`, `SAT I Superscore (CR + M) - Verified`, `SAT R Superscore`) %>%
  mutate(Accepted = 0)

colnames(denied) <- c("Ref", "Round", "Rigor","GPA","ACT","SAT_I", "SAT_R","Accepted")
```


Combine the both the admitted and denied student data into a single data frame.  

```{r}
applied_raw <- rbind(admit,denied)
applied <- applied_raw
```


Remove any gpa over 5.0 or equal to zero.  

```{r}
applied[applied$GPA > 5 | applied$GPA == 0,"GPA"] <- NA
```


Convert the old SAT scores (SAT_I) to the new scores (SAT_R) using the conversion chart.  

```{r}
SAT_conversions <- fread("../data/old_to_new_SAT.csv", sep = ",")

converted_SAT_I <- left_join(applied[, c("Ref","SAT_I","SAT_R")],SAT_conversions, by = "SAT_I")[,"SAT_R.y"]
applied <- cbind(applied, converted_SAT_I)

```


Pick the max score from SAT_R and the converted SAT_I scores.  

```{r}

applied[is.na(applied)] <- -1
SAT_final <- apply(applied[,c("SAT_R","converted_SAT_I")],1,max)
applied <- cbind(applied, SAT = SAT_final)
applied[applied == -1] <- NA

```


Remove all other SAT columns except for SAT_final, and write to CSV file for later analysis.  

```{r}
applied_clean <- applied[,c(1:5,10,8)]
fwrite(applied_clean, "../data/applied_clean.csv")
```


Build logistic regression models for predicting acceptance based on combinations of GPA, ACT, and SAT scores.  

```{r}
data <- applied_clean
#Columns
accepted <- data$Accepted
course_selection_rigor <- data$Rigor
gpa <- data$GPA
sat <- data$SAT
act <- data$ACT
essay <- data$Essay
```


```{r regression}
model_both <- glm(accepted ~ gpa + sat + act, 
             data = data, 
             family = binomial)
coefficients(model_both)

model_act <- glm(accepted ~ gpa + act, 
             data = data, 
             family = binomial)
coefficients(model_act)

model_sat <- glm(accepted ~ gpa + sat, 
             data = data, 
             family = binomial)
coefficients(model_sat)
```


Create bins for GPA and get counts for each bin.  

```{r}
admit_clean <- applied_clean %>%
  filter(Accepted == 1)
GPA_bin_labels = c("0 - 0.5",
                   "0.51 - 1.0",
                   "1.01 - 1.5",
                   "1.51 - 2.0",
                   "2.01 - 2.5",
                   "2.51 - 3.0",
                   "3.01 - 3.5",
                   "3.51 - 4.0",
                   "4.01 - 4.5",
                   "4.51 - 5.0")

GPA_binned <- cut(admit_clean$GPA, breaks = seq(0,5,by = 0.5), labels = GPA_bin_labels, right = FALSE, include.lowest = TRUE)
counts_for_GPA_bins <- summary(GPA_binned)
counts_for_GPA_bins
```


Create bins for SAT and get counts for each bin.  

```{r}
SAT_bin_labels = c("400 - 500",
                   "510 - 600",
                   "610 - 700",
                   "710 - 800",
                   "810 - 900",
                   "910 - 1000",
                   "1010 - 1100",
                   "1110 - 1200",
                   "1210 - 1300",
                   "1310 - 1400",
                   "1410 - 1500",
                   "1510 - 1600")

SAT_binned <- cut(admit_clean$SAT, breaks = seq(400,1600,by = 100), labels = SAT_bin_labels, right = FALSE, include.lowest = TRUE)
counts_for_SAT_bins <- summary(SAT_binned)
counts_for_SAT_bins
```


Create bins for ACT and get counts for each bin. 

```{r}
ACT_bin_labels = c("0 - 3",
                   "4 - 6",
                   "7 - 9",
                   "10 - 12",
                   "13 - 15",
                   "16 - 18",
                   "19 - 21",
                   "22 - 24",
                   "25 - 27",
                   "28 - 30",
                   "31 - 33",
                   "33 - 36")

ACT_binned <- cut(admit_clean$ACT, breaks = seq(0,36,by = 3), labels = ACT_bin_labels, right = FALSE, include.lowest = TRUE)
counts_for_ACT_bins <- summary(ACT_binned)
counts_for_ACT_bins
```

