---
title: "Getting Coefficients"
author: "Adriana Souza"
date: "April 14, 2018"
output: pdf_document
---

# Cleaning

1. Added variable "Accepted" to "Admitted Students" and "Denied Students" -- "1" for admitted students, "0" for denied students.
2. Merged "Course Selection Rating" with "2017 Course Sel rating" into "Rating" column for "Denied Students" data set.
3. Merged "Final NCF GPA - 2017", "Final NCF GPA - 2016 and prior", "2017 GPA Calc", and "NCF GPA", into one "GPA"" column. 
4. Merged "Essay Rating - 2017" and "Essay Rating - 2017" into one "Essay" columns for the "Denied Students" data. 

```{r data}
#Load data
data <- read.csv('../data/applied_clean.csv')

#Columns
accepted <- data$Accepted
course_selection_rigor <- data$Rigor
gpa <- data$GPA
gpa2 <- gpa
gpa2[gpa2 > 5] <- NA
gpa <- gpa2
sat <- data$SAT
act <- data$ACT
essay <- data$Essay

```


```{r regression}
model_both <- glm(accepted ~ gpa + sat + act, 
             data = data, 
             family = binomial)
coef(model_both)

model_act <- glm(accepted ~ gpa + act, 
             data = data, 
             family = binomial)
coef(model_act)

model_sat <- glm(accepted ~ gpa + sat, 
             data = data, 
             family = binomial)
coef(model_sat)
```

